---
title: "DATA 607 Week 9 Assignment : Web APIs"
author: "Fan Xu"
date: "10/27/2019"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# I. Assignment Overview

The New York Times web site provides a rich set of APIs, as described here: <https://developer.nytimes.com/apis>(https://developer.nytimes.com/apis) 

You’ll need to start by signing up for an API key. 

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.



# II. Load Package
```{r 0, message=FALSE, warning=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(stringr)
library(kableExtra)
library(ggplot2)
library(forcats)
```


```{r 1, echo=FALSE}
api_key <- 'HzMYhjtkJ1eZ8xJC98d9Qp0rsABz8lU3'

```


# III. Set filter
I will use Article Search API to look up articles. The date range is the whole year 2019. Search results from the first 5 pages will be downloaded.
```{r 2}
base_url = 'https://api.nytimes.com/svc/search/v2/articlesearch.json'
start_date <- "20190101"
end_date <- "20191231"
first_page = 1
last_page = 5

```


# IV. Read data and Convert to DataFrame
## a. Create a dataframe to consolidate outputs from all pagination in scope.
```{r 3}
df_json <- data.frame(
         pub_date=character(),
         headline=character(),
         abstract=character(),
         lead_paragraph=character(),
         author=character(),
         type_of_material=character(),
         source=character(),
         section_name=character(),
         word_count=integer(),
         web_url=character(),
         stringsAsFactors = FALSE
)
```

## b. Get data, perfrom tidying and produce final output
a. set a `for` loop to get contents from the first 5 pages of search result. `api_key` is hidden.

b. `GET` function from package `httr` is used to get data from data source.

c. `fromJSON` function from package `jsonlite` is used to convert json files to data frames.

d. store data from each search page to data frame `df_temp`.

e. consolidate `df_temp` from each search page into data frame `df_json` as final output.
```{r 4}

for (page in first_page:last_page){
url = str_c(base_url, 
            '?begin_date=',start_date,
            '&end_date=',end_date,
            '&offset=',page,
            '&api-key=',api_key,
            sep = '')

articles <- GET(url)

#http_status(articles)

stop_for_status(articles)

df_temp <- fromJSON(url) %>%
 .$response %>%
 .$doc %>%
  mutate(headline = headline$main) %>%
  mutate(byline = byline$original) %>%
  mutate(byline = str_remove(byline, '^By ')) %>%
  mutate(pub_date = as.Date(pub_date)) %>%
  rename(author = byline) %>%
  select(pub_date,
         headline,
         abstract,
         lead_paragraph,
         author,
         type_of_material,
         source,
         section_name,
         word_count,
         web_url)
  

df_json <- rbind(df_json,df_temp)
#print(c("Current Page: ", page))
closeAllConnections()
Sys.sleep(30)

df_json
}
```


```{r 5}
df_json %>%
  kable() %>% 
  kable_styling() %>%
  scroll_box(width = "810px", height = '500px')
```

# V. Extra: Plots for Analysis
```{r 6}
ggplot(data=df_json, aes(x=fct_rev(fct_infreq(section_name)), y=..count.., fill = ..count..))+
  geom_bar(stat='count')+
  geom_text(stat = 'count', aes(label=..count..),size = 3, color= 'white',position = position_stack(vjust = 0.8))+
  scale_fill_gradient(low = 'deeppink4', high = 'deeppink1')+
  xlab('section name')+
  ggtitle('Number of Articles by Section Name')+
  coord_flip()

ggplot(data=df_json, aes(x=fct_rev(fct_infreq(type_of_material)), y=..count.., fill = ..count..))+
  geom_bar(stat='count')+
  geom_text(stat = 'count', aes(label=..count..),size = 3, color= 'white',position = position_stack(vjust = 0.8))+
  scale_fill_gradient(low = 'tan4', high = 'tan1')+
  xlab('type of material')+
  ggtitle('Number of Articles by Type of Material')+
  coord_flip()

ggplot(data=df_json, aes(x=word_count, fill=..count..))+
  geom_histogram(bins=30,color="black")+
  scale_fill_gradient(low = 'cyan4', high = 'cyan1')+
  ggtitle('Histogram: Word Count')
  
  
```